---
title: "Lists"
author: "Sandra Alvarez-Carretero"
date: "February 26, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The next vector that we are going to work with is the list. The main function of the lists is to keep everything ordered under a specific name. You can even think of it as a dictionary in Python, because you can access the elements of a list by the name you have given to every entry. In order to understand this better, let's see an example!

## 1. Some examples
Susan and Luisa have collected some fossils, so they have decided to identify them. In order to differentiate which species have been sampled by who, we have created the following list:

```{r}
fossils <- list( susan = c( "Tomarctus", "Smilodon", "Hesperocyon" ),
                 luisa = c( "Aelurodon", "Enhydrocyon" ) )
```

If you want to know which species have been sampled by Susan and which of them by Luisa, the syntax to access the entries of a list is to first type the name of the list, followed by the dollar sign `$`, and then the name of the entry (in this case either `susan` or `luisa`). However, if you know the position of the entry you want to retrieve, you can also type the name of the list and then, in **DOUBLE** square brackets, the number corresponding to the position of the entry you are interested to retrieve:

```{r}
fossils$susan
all.equal( fossils$susan, fossils[[ 1 ]] )

fossils$luisa
all.equal( fossils$luisa, fossils[[ 2 ]] )

```

As you can see, it is much understandable to use the dollar sign followed by the name of the entry, so we recommend you to name the entries of your lists!
You can also access specific elements of each entry using the single square brackets, as we did when we wanted to access different elements of a vector. If you want to get the first element of the entry named `susan`, just type:

```{r}
fossils$susan[ 1 ]
all.equal( fossils[[ 1 ]][ 1 ], fossils$susan[ 1 ] )
```

Maybe, you want to change the name of the entries at some point. If this is the case, you can use the function `names()`. For instance, imagine that now we want to have the first letter of the names of the entries as a capital letter because it is a name and you are very picky about grammar:

```{r}
names( fossils ) <- c( "Susan", "Luisa" )
fossils$Susan
fossils$Luisa 
```

Furthermore, there is a function called `attach()` that lets you have your list as part of the R path, so you do not need to type `fossils$Susan` or `fossils$Luisa` to retrieve the elements of each entry. You should just type `Susan` or `Luisa`, the names of the entry.

### **IMPORTANT!!**
When you finish using your list, please **DETACH** the list from the path. This might save you from more than one headache in the future!

```{r} 
attach( fossils )
Susan
Luisa 

detach( fossils )
fossils$Susan
fossils$Luisa
```

## 2. Exercises

### Exercise 1  
Create a list with the favourite colours of 3 people.  
***HINT***: Think how many entries you need for that!

### Exercise 2  
Access the second entry of your list in the two ways you have learnt and check they return the same value!

### Exercise 3  
Access the first and second elements of the third entry of your list

### Exercise 4  
Find which are the colours that both the person in the first and in the second entry like.
***HINT***: Remember some functions we used with the numeric vectors!


