---
title: "Factors"
author: "Sandra Alvarez-Carretero"
date: "February 26, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

If you want to arrange your data into levels, it is important that you convert your data into factor class. Learn how in this tutorial!

## 1. Some examples

Imagine that you have a big database where the information people have provided you with can be classified into three categories: "Academia", "Industry", "Both". You have asked 10 people about the places that they work. The following vector contains the answers of these people:

```{r}
work <- c( "Academia", "Academia", "Academia", "Industry", "Both",
           "Both", "Both", "Industry", "Both", "Academia", "Industry")
```

In this case, you see that these three answers are repeated, which means that you can think of them as `levels` in an object of class factor in R. In order to identify the levels of a vector, you just have to type the following command using the function `factor()`:

```{r}
types_work <- factor( work )
types_work
class( types_work )
```

When you print the content of `types_work`, you see that the 10 names appeared first, labelled from 1 to 10. After that, you see that it is printed `3 Levels: Academia ... Industry`. This means that your `types_work` variable contains the levels of your data already!
If you want to retrieve only the name of your `levels`, you can use the function `levels()`:
```{r}
levels( types_work )
```

One of the most important utilities of the factors is the way they can be used to plot the information they refer to. Let's add now more information that we had already included in our database:   

   a) The amount of time they had worked in a previous job   
   b) The amount of time they have been working in current job   

```{r}
time_previous <- c( 0, 1, 0, 5, 6, 10, 2, 10, 20, 1 )
time_current <- c( 5, 8, 3, 2, 5, 3, 3, 1, 2, 4 )
```

Now, let's plot the content of these variables!

```{r}
# Plot without information about level
plot( time_previous, time_current )

# Plot with information about level
plot( time_previous, time_current, pch = as.integer( types_work ) )

```

Note that, in the second plot, you can see different symbols. This is because the argument `pch` assigns a different symbol to each level so you can identify which point belongs to which level. However, you also need the legend to keep track of this! 
Run the following code by selecting the code within lines 64 to 67 and then pressing `ctrl+RETURN` or `command+RETURN` (depending on Windows or Mac users) to run the whole code at once:

```R
plot( time_previous, time_current, pch = as.integer( types_work ) )
legend( "topright",
         legend = levels( types_work ),
         pch = unique( as.integer( types_work ) ) )
```

***NOTE***: Read the documentation of `?unique`.

## 2. Exercise
Partner with some of your class mates and discuss all the examples presented in this tutorial. Change parameters, plot different values, look into new functions, etc. For instance, you could try to see how you can have one specific colour in the plot for each level!

Run the following code and try to understand what the function `locator()` does an what we are doing here: 
```R
plot( time_previous, time_current, pch = as.integer( types_work ) )
points <- locator( n = 1 )
legend( x = points$x, y = points$y,
       legend = levels( types_work ),
       pch = unique( as.integer( types_work ) ) )
```


